Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-24 manuscript/paper.tex
highlighted points resolved

The results in Eqs.~\ref{eq:theta0-Es},\ref{eq:theta0-E0}
are summarised in Fig.~\ref{fig:t2g-alpha0},
where $E_{xy}$, $E_{yz/zx}$ and $E_0^{t_{2g}}$ are shown
as a function of $\beta$
at three different values of $q_A/q_o=1/2,1,2$.
We have 
assumed $D_2=\sqrt{D_4}$ and
taken $\Delta_o=4q_o/15a$ throughout this paper,
which fixes the value of $D_4=1/200$, 
and corresponds to $\Delta= 2.073$ eV at 
$q_o=2$ e and $a=7.0$ bohr.
We present the energies as a percentage of $\Delta_o$
by calculating all quantities at the above values of $q_o,a$.
Fig.~\ref{fig:t2g-alpha0} shows that
the the competition between A and B cages
determines the ground state and the size of the splitting between the 
$t_{2g}$ states
$E_{xy}$, and $E_{yz/zx}$.
In Fig.~\ref{fig:t2g-alpha0}(a),
where $q_A/q_o=1/2$ ($q_B/q_o=5/2$),
the effect of B cage dominates.
Since it's negative (see Eqs.~\ref{eq:theta0-delB2},\ref{eq:theta0-delB4}),
it reverses the order in Eq.~\ref{eq:theta0-Es}
and $E_{xy}$ becomes lower than $E_{yz/zx}$, by about $6\%$ of $\Delta_o$ at $15^\circ$.
When A cage dominates at large $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-alpha0}(c) for $q_A/q_o=2$,
$E_{yz/zx}$ are the single particle ground states (again with a splitting of $\sim6\%$ of $\Delta_o$ at $15^\circ$),
whereas the net effect of both cages
is reduced in Fig.~\ref{fig:t2g-alpha0}(b)
that corresponds to $q_A/q_o=1$
and we only obtain a splitting of $\sim2\%$ of $\Delta_o$ at $15^\circ$
between $E_{xy}$ and $E_{yz/zx}$ levels.
Obviously, there will also be a point near 
$q_A/q_o=1$ where A and B completely 
cancel each others effect (not shown in Fig.~\ref{fig:t2g-alpha0}).
We also see that $E_0^{t_{2g}}$ changes from a slight negative shift to a slight positive shift
with an increase in $q_A/q_o$ as we
go from Fig.~\ref{fig:t2g-alpha0}(a)
to Fig.~\ref{fig:t2g-alpha0}(c).
This is because the rotation suppresses the
effect of AB cages on the
gap $\Delta$ between $e_{g}$ and $t_{2g}$.
Since A supports O while B opposes it,
$\Delta$ slightly widens at small $q_A$ and shrinks at large $q_A$,
shifting $E_0^{t_{2g}}$ slightly down or up with $\beta$.

@@

\az{

Summarise t2g results here.\\

copy from summary section...??

tilt vs rotation, preference:\\

>>>>

The results in Eqs.~\ref{eq:theta0-Es},\ref{eq:theta0-E0}
are summarised in Fig.~\ref{fig:t2g-alpha0},
where $E_{xy}$, $E_{yz/zx}$ and $E_0^{t_{2g}}$ are shown
as a function of $\beta$
at three different values of $q_A/q_o=1/2,1,2$.
We have 
assumed $D_2=\sqrt{D_4}$ and
taken $\Delta_o=4q_o/15a$ throughout this paper,
which fixes the value of $D_4=1/200$, 
and corresponds to $\Delta_o= 2.073$ eV at 
$q_o=2$ e and $a=7.0$ bohr.
We present the energies as a percentage of $\Delta_o$
by calculating all quantities at the above values of $q_o,a$.
Fig.~\ref{fig:t2g-alpha0} shows that
the the competition between A and B cages
determines the ground state and the size of the splitting between the 
$t_{2g}$ states
$E_{xy}$, and $E_{yz/zx}$.
In Fig.~\ref{fig:t2g-alpha0}(a),
where $q_A/q_o=1/2$ ($q_B/q_o=5/2$),
the effect of B cage dominates.
Since it's negative (see Eqs.~\ref{eq:theta0-delB2},\ref{eq:theta0-delB4}),
it reverses the order in Eq.~\ref{eq:theta0-Es}
and $E_{xy}$ becomes lower than $E_{yz/zx}$, by about $6\%$ of $\Delta_o$ at $15^\circ$.
When A cage dominates at large $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-alpha0}(c) for $q_A/q_o=2$,
$E_{yz/zx}$ are the single particle ground states (again with a splitting of $\sim6\%$ of $\Delta_o$ at $15^\circ$),
whereas the net effect of both cages
is reduced in Fig.~\ref{fig:t2g-alpha0}(b)
that corresponds to $q_A/q_o=1$
and we only obtain a splitting of $\sim2\%$ of $\Delta_o$ at $15^\circ$
between $E_{xy}$ and $E_{yz/zx}$ levels.
Obviously, there will also be a point near 
$q_A/q_o=1$ where A and B completely 
cancel each others effect (not shown in Fig.~\ref{fig:t2g-alpha0}).
We also see that $E_0^{t_{2g}}$ changes from a slight negative shift to a slight positive shift
with an increase in $q_A/q_o$ as we
go from Fig.~\ref{fig:t2g-alpha0}(a)
to Fig.~\ref{fig:t2g-alpha0}(c).
This is because the rotation suppresses the
effect of AB cages on the
gap $\Delta$ between $e_{g}$ and $t_{2g}$.
Since A supports O while B opposes it,
$\Delta$ slightly widens at small $q_A$ and shrinks at large $q_A$,
shifting $E_0^{t_{2g}}$ slightly down or up with $\beta$.

@@

\com{
Summarise t2g results here.\\
copy from summary section...??
tilt vs rotation, preference:\\
 }
